---
title: "Homework 3 - Jacob Richards"
output:
  pdf_document: default
  html_document: default
date: "2024-09-25"
---
1) Develop an S4 class that implements the CUSUM chart including a show()
method. You will need to store the lower and upper signals as slots.
Hint: See example in charts.R in shewhart.chart.zip.


```{r}


CL = spc::xcusum.crit(k = 0.5, L0 = 370.0, mu0 = 0, hs = 0, sided = "two", r = 30)


# Define the class
setClass("Data_stream", 
         representation(delta = "numeric",
                        x = "numeric",
                        random_t = "numeric",
                        Upper_signal = "numeric",
                        Lower_signal = "numeric",
                        IC.ARL  = "numeric",
                        CL      = "numeric",
                        time = "numeric",  # Add this line to include 'time' slot
                        time_values = "numeric",
                        stopped = "logical",
                        x_values = "numeric",
                        Upper_signal_vector = "numeric",  # Stores all upper signals
                        Lower_signal_vector = "numeric"))  # Stores all lower signals


```

```{r}
setGeneric("create_Data_Stream", 
           function(IC.ARL = 370, random_t = as.integer(runif(1, 0, 100)), 
             time = 0, delta = 1, CL = spc::xcusum.crit(k = 0.5, L0 = 370.0, mu0 = 0,
             hs = 0, sided = "two", r = 30)) {
             standardGeneric("create_Data_Stream")
           })

setMethod("create_Data_Stream", signature = "ANY", 
          function(IC.ARL = 370, random_t = as.integer(runif(1, 0, 100)), 
                   time = 0, delta = 1, CL = spc::xcusum.crit(k = 0.5, L0 = 370.0,
                   mu0 = 0, hs = 0, sided = "two", r = 30)) {

    new_data <- new("Data_stream", 
                IC.ARL = IC.ARL, random_t = random_t, delta = delta,
                x = 1, Upper_signal = 0, Lower_signal = 0, 
                CL = spc::xcusum.crit(k = 0.5, L0 = 370.0, mu0 = 0, hs = 0, sided = "two",
                r = 30), time = time, stopped = FALSE,
                x_values = numeric(),
                Upper_signal_vector = numeric(), 
                Lower_signal_vector = numeric(),  
                time_values = numeric())  

    time_counter <- 0  

    while (time_counter < random_t) {
      new_data@x <- rnorm(1, 0, 1)  # Generate data from N(0, 1) before random_t
      new_data@x_values <- c(new_data@x_values, new_data@x)  # Store x

      # Update the Upper and Lower CUSUM signals
      new_data@Lower_signal <- min(0, new_data@Lower_signal + new_data@x + 0.5)
      new_data@Upper_signal <- max(0, new_data@Upper_signal + new_data@x - 0.5)

      # Store the updated signals and time
      new_data@Upper_signal_vector <- c(new_data@Upper_signal_vector, new_data@Upper_signal)
      new_data@Lower_signal_vector <- c(new_data@Lower_signal_vector, new_data@Lower_signal)
      new_data@time_values <- c(new_data@time_values, time_counter)  # Store the time step

      # Update the data stream and check for stopping condition
      new_data <- update_Data_Stream(new_data, new_data@x)

      # Break out of the loop if stopped
      if (new_data@stopped) {
        break
      }
      
      time_counter <- time_counter + 1  # Increment the time counter
    }

    # Post-change phase
    while (!new_data@stopped) {
      new_data@x <- rnorm(1, delta, 1)  # Generate data from N(delta, 1) after random_t
      new_data@x_values <- c(new_data@x_values, new_data@x)  # Store x

      # Update the Upper and Lower CUSUM signals
      new_data@Lower_signal <- min(0, new_data@Lower_signal + new_data@x + 0.5)
      new_data@Upper_signal <- max(0, new_data@Upper_signal + new_data@x - 0.5)

      # Store the updated signals and time
      new_data@Upper_signal_vector <- c(new_data@Upper_signal_vector, new_data@Upper_signal)
      new_data@Lower_signal_vector <- c(new_data@Lower_signal_vector, new_data@Lower_signal)
      new_data@time_values <- c(new_data@time_values, time_counter)  # Store the time step

      # Update the data stream and check for stopping condition
      new_data <- update_Data_Stream(new_data, new_data@x)

      
      if (new_data@stopped) {
        break
      }

      time_counter <- time_counter + 1  # Increment the time counter
    }

    return(new_data)
})


```




```{r}
setGeneric("update_Data_Stream", 
           function(new_data, x) {
             standardGeneric("update_Data_Stream")
           })
setMethod("update_Data_Stream", signature = "Data_stream", 
          function(new_data, x) {
  # Increment the time step
  new_data@time <- new_data@time + 1
  
  # Check if the last values of Upper and Lower signal vectors exceed the control limit
  if (length(new_data@Upper_signal_vector) > 0) {
    new_data@stopped <- (tail(new_data@Upper_signal_vector, 1) > new_data@CL) || 
                        ((tail(new_data@Lower_signal_vector, 1)) < -new_data@CL)
  }
  if (length(new_data@Lower_signal_vector) > 0 ) {
    new_data@stopped <- (tail(new_data@Upper_signal_vector, 1) > new_data@CL) || 
                        ((tail(new_data@Lower_signal_vector, 1)) < -new_data@CL)
  }

  
  
  return(new_data)
})


```


```{r}
setMethod("show", signature = "Data_stream", function(object) {
  # Plot the upper signal with a blue line
  plot(object@time_values, object@Upper_signal_vector, type = "l", col = "blue", 
       ylab = "Signal", xlab = "Time", main = "Upper and Lower Signals Over Time", 
       ylim = range(c(object@Upper_signal_vector, object@Lower_signal_vector, object@CL,-object@CL)))  
  
  # Add the lower signal with a red line
  lines(object@time_values, object@Lower_signal_vector, col = "red")  # Use lines() for lower signal
  
  object@CL = spc::xcusum.crit(k = 0.5, L0 = 370.0, mu0 = 0, hs = 0, sided = "two", r = 30)
  
  # Add horizontal lines for control limits
  abline(h = object@CL, col = "green", lty = 2)  # Positive control limit
  abline(h = -object@CL, col = "orange", lty = 2)  # Negative control limit
  abline(v = object@random_t, col = "pink", lty = 2)
  
cat("Data steam display \n")
            
            cat("control limit reached at this time  =", object@time, "\n")
            cat("mean changed at this time  =", object@random_t, "\n")
            cat("absolute value control limit", object@CL, "\n")
            cat("Last five upper limits:", tail(object@Upper_signal_vector, 5), "\n")
            cat("Last five lower limits:", tail(object@Lower_signal_vector, 5), "\n")
            
})


```


It is unlikely but sometimes the control limit will be reached before the mean changes. 
```{r}

new_data <- create_Data_Stream()

new_data <- update_Data_Stream(new_data, new_data@x)  

show(new_data)

```






Let k = 0.5 and CL = spc::xcusum.crit(k = 0.5, L0 = 370.0, mu0 = 0, hs = 0,
sided = "two", r = 30). Generate N = 10,000 i.i.d. standard normal data
streams and compute the associated stopping times CUSUM.
a) Plot a histogram for the resulting 10,000 CUSUM values.
b) Calculate the sample mean of CUSUM (referred to as “in-control average
run length (ARL)”). The value should be close to 370.2)



```{r}

library(spc)

CL <- spc::xcusum.crit(k = 0.5, L0 = 370, mu0 = 0, hs = 0, sided = "two", r = 30)

# Define the class
setClass("Data_stream_2", 
         representation(delta = "numeric",
                        x = "numeric",
                        random_t = "numeric",
                        Upper_signal = "numeric",
                        Lower_signal = "numeric",
                        CL = "numeric",
                        stopped = "logical",
                        x_values = "numeric",
                        Upper_signal_vector = "numeric",  
                        Lower_signal_vector = "numeric",  
                        iteration_count = "numeric"))     

# Define create_Data_Stream_2 method
setGeneric("create_Data_Stream_2", 
           function(random_t = as.integer(runif(1, 0, 100)), 
                    delta = 0, CL = spc::xcusum.crit(k = 0.5, L0 = 370.0, mu0 = 0,
                    hs = 0, sided = "two", r = 30),k=0.5, l0=370,mu0=0, hs=0, sided="two", r=30)
             standardGeneric("create_Data_Stream_2"))

setMethod("create_Data_Stream_2", signature = "ANY", 
          function(random_t = as.integer(runif(1, 0, 100)), 
                    delta = 0, CL = spc::xcusum.crit(k = 0.5, L0 = 370.0, mu0 = 0,
                    hs = 0, sided = "two", r = 30), k=0.5, l0=370,mu0=0, hs=0, sided="two", r=30) {

   

    # Initialize a new Data_stream_2 object
    new_data_2 <- new("Data_stream_2", 
                      delta = delta,
                      x = 1, 
                      Upper_signal = 0, 
                      Lower_signal = 0, 
                      CL = CL, 
                      stopped = FALSE,
                      x_values = numeric(),
                      Upper_signal_vector = numeric(),  
                      Lower_signal_vector = numeric(),  
                      iteration_count = 0)  

    while (!new_data_2@stopped) {
      # Generate data from N(0, 1)
      new_data_2@x <- rnorm(1, 0, 1)
      
      # Store x values
      new_data_2@x_values <- c(new_data_2@x_values, new_data_2@x)

      # Update the Upper and Lower CUSUM signals
      new_data_2@Lower_signal <- min(0, new_data_2@Lower_signal + new_data_2@x + k)
      new_data_2@Upper_signal <- max(0, new_data_2@Upper_signal + new_data_2@x - k)

      # Store the updated signals
      new_data_2@Upper_signal_vector <- c(new_data_2@Upper_signal_vector, new_data_2@Upper_signal)
      new_data_2@Lower_signal_vector <- c(new_data_2@Lower_signal_vector, new_data_2@Lower_signal)

      # Increment the iteration count
      new_data_2@iteration_count <- new_data_2@iteration_count + 1
      
      # Update the data stream and check for stopping condition
      new_data_2 <- update_Data_Stream_2(new_data_2, new_data_2@x)

      if (new_data_2@stopped) {
        break
      }
    }

    return(new_data_2)
})

# Define update_Data_Stream_2 method
setGeneric("update_Data_Stream_2", 
           function(new_data_2, x_2) {
             standardGeneric("update_Data_Stream_2")
           })

setMethod("update_Data_Stream_2", signature = "Data_stream_2",function(new_data_2, x_2) {
  
  # Check if the last values of Upper and Lower signal vectors exceed the control limit
  if (length(new_data_2@Upper_signal_vector) > 0) {
    new_data_2@stopped <- (tail(new_data_2@Upper_signal_vector, 1) > new_data_2@CL) || 
                        (tail(new_data_2@Lower_signal_vector, 1) < -new_data_2@CL)
  }
            
            if (length(new_data_2@Lower_signal_vector) > 0) {
    new_data_2@stopped <- (tail(new_data_2@Upper_signal_vector, 1) > new_data_2@CL) || 
                        (tail(new_data_2@Lower_signal_vector, 1) < -new_data_2@CL)
  }

  return(new_data_2)
})

# Run and collect stopping times for N = 10,000 streams
N <- 10000
stopping_times <- numeric(N)

for (i in 1:N) {
  new_data_2 <- create_Data_Stream_2()  # CL will be dynamically calculated
  stopping_times[i] <- new_data_2@iteration_count
}
summary(stopping_times)
hist(stopping_times, breaks = 50, main = "Histogram of Stopping Times",
     xlab = "Stopping Time", col = "lightblue")
print(mean(stopping_times))


```




